---
title: "001_EPA_explore"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load data
```{r setup, include=FALSE}
#Load data

# nla_2017 <- read.csv(here("data/NLA/data/nla_2017_water_chemistry_chla-data.txt"))
# nla_2012 <- read.csv(here("data/NLA/data/nla2012_waterchem_wide.txt"))
# nla_2007 <- read.csv(here("data/NLA/data/nla2007_chemical_conditionestimates_20091123.txt"))
# 
# glimpse(nla_2007)
# 
# nla_2007_trim <- nla_2007 %>%
#   select(SITE_ID, VISIT_NO, SITE_TYPE, TNT,
#          LAT_DD, LON_DD, ST, NUTREG_NAME,
#          ECO3_X_ORIGIN, HUC_2, HUC_8,
#          PTL:COND, CHLA) %>%
#   rename(TP_ugL=PTL,
#          TN_ugL=NTL,
#          DOC_mgL=DOC)


source_folder <- nla_get(2012)
dt_2012 <- nla_load(2012, source_folder = source_folder)

source_folder2 <- nla_get(2007)
dt_2007 <- nla_load(2007, source_folder = source_folder2)


waterchem_2007 <- dt_2007$WaterQuality
waterchem_2012 <- dt_2012$waterchem_wide
siteinfo_2012 <- dt_2012$wide_siteinfo
chla_2012 <- dt_2012$chla_wide

glimpse(waterchem_2007)
glimpse(waterchem_2012)
glimpse(siteinfo_2012)
glimpse(chla_2012)


##Compile 2012 data
siteinfo_2012_western <- siteinfo_2012 %>%
  filter(STATE %in% c("CA", "UT", "NV",
                      "WA", "OR", "ID",
                      "MT", "WY", "CO",
                      "NM", "AZ")) %>%
  select(SITE_ID, SITEID_07, VISIT_NO,
         UID, DATE_COL, AREA_HA, CAT_UNIQUE,
         EPA_REG, HUC2, HUC8, LAKE_ORIGIN, LAKE_ORIGIN12,
         LAT_DD83, LON_DD83, MAJ_BASIN, SIZE_CLASS)

waterchem_2012_trim <- waterchem_2012 %>%
  select(UID, DOC_RESULT, NITRATE_N_RESULT,
         NTL_RESULT, PTL_RESULT, TURB_RESULT)

chla_2012_trim<- chla_2012 %>%
  select(UID, CHLX_RESULT)


nla2012 <- left_join(siteinfo_2012_western, waterchem_2012_trim, by="UID")
nla2012 <- left_join(nla2012, chla_2012_trim, by="UID")

glimpse(nla2012)

###Compile 2007 data
waterchem_2007_trim <- waterchem_2007 %>%
  dplyr::select(SITE_ID, DATE_COL, DOC, NTL, PTL,
         TURB, NO3, CHLA, SECMEAN, NO3N_PPM)
chemconditions_2007 <- dt_2007$Chemical_ConditionEstimates 
# glimpse(chemconditions_2007)
chemconditions_2007_trim <- chemconditions_2007 %>%
  dplyr::select(SITE_ID, LAT_DD, LON_DD, LAKE_ORIGIN, HUC_2,
         HUC_8, PTL, NTL, DOC, TURB, CHLA) %>%
  rename(PTL_2=PTL,
         NTL_2=NTL,
         DOC_2=DOC,
         TURB_2=TURB,
         CHLA_2=CHLA)
siteinfo_2007<- dt_2007$SampledLakeInformation
glimpse(siteinfo_2007)
siteinfo_2007_trim <- siteinfo_2007 %>%
  dplyr::select(SITE_ID, ST, DATE_COL, NHDNAME, AREA_HA, SIZE_CLASS, LAKEAREA,
         LAKEPERIM, DEPTHMAX, DEPTH_X, ELEV_PT) %>%
    filter(ST %in% c("CA", "UT", "NV",
                      "WA", "OR", "ID",
                      "MT", "WY", "CO",
                      "NM", "AZ")) 

nla2007<- left_join(siteinfo_2007_trim, chemconditions_2007_trim, by="SITE_ID")
nla2007<- left_join(nla2007,waterchem_2007_trim, by=c("SITE_ID","DATE_COL"))
```


#Maps - 2007 data
##Basic distribution, size, and lake types
```{r}
#Map 2012 western NLA sites

northamerica<- ne_countries(country = "united states of america", returnclass= "sf")
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-130, -100),
           ylim = c(30, 50)) +
  geom_point(data=nla2007,aes(LON_DD,LAT_DD,fill=factor(HUC_2)),
             shape=21, color="black", size=3)+
  scale_fill_discrete_diverging(rev=TRUE,
                                 labels=basinNames)+
  ggtitle("2007 Lakes Sampled")

# ggplot(data = northamerica) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(
#     xlim = c(-130, -100),
#     ylim = c(30, 50)) +
#   geom_point(data=nla2012,aes(LON_DD83,LAT_DD83,fill=SIZE_CLASS),
#              shape=21, color="black")+
#   scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)


#How many of these lakes are reservoirs?  
nla2007 %>%
  select(HUC_2,LAKE_ORIGIN) %>%
  # filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL"))%>%
  group_by(HUC_2,LAKE_ORIGIN)%>%
  mutate(num_lakes=n()) %>%
  distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=num_lakes, fill=factor(HUC_2)))+
  geom_bar(stat="identity",  color="black")+
  ylab("Number of lakes")+
  xlab("Lake type")+
  facet_wrap(.~factor(HUC_2), scales="free_y",
              labeller = labeller(name = basin.Names))+
  scale_fill_discrete_diverging(rev=TRUE,
                                labels=basinNames)+
  ggtitle("Lake origin by major basin")

#Lakes of different size classes
nla2007 %>%
  select(SIZE_CLASS) %>%
  filter(!SIZE_CLASS=="")%>%
  # pivot_longer(-1) %>%
  group_by(SIZE_CLASS) %>%
  mutate(num_lakes=n()) %>%
  distinct() %>% #remove duplicates
  ggplot(aes(x=SIZE_CLASS, y=num_lakes, fill=SIZE_CLASS))+
  geom_bar(stat="identity",  color="black")+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Number of lakes")+
  xlab("Size class")+
  # geom_text(aes(label=load_calc_possible_n), vjust=-1) +
  theme(legend.position="none")+
  ggtitle("2007 - Number of lakes by size class")

```


##Chemistry
###Chlorophyll by lake type, size, basin
```{r}
#Chlorophyll across lakes with data
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2007 %>%
               filter(LAKE_ORIGIN %in% c("MAN-MADE","NATURAL")),
                                  aes(LON_DD,LAT_DD,
                                   fill=log10(CHLA),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Greens", rev=TRUE)+
  scale_shape_manual(values=c(21,23))

#Chlorophyll in lakes of different size classes
#Lakes of different size classes
nla2007 %>%
  select(SIZE_CLASS, CHLA, LAKE_ORIGIN) %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(CHLA), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 chlorophyll a")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Greens", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)+
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) 



#Chlorophyll in lakes of different major basins
nla2007 %>%
  select(HUC_2, CHLA,LAKE_ORIGIN) %>%
  filter(!HUC_2 %in% c(""))%>%
  mutate(HUC_2=factor(HUC_2,
                      labels=c("Missouri","Ark-Red-White", "Rio Grande",
                 "Upper Colorado","Lower Colorado","Great Basin","Pacific Northwest",
                 "California"))) %>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(CHLA), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  ylab("Log10 chlorophyll a")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Terrain2", rev=TRUE)+
  facet_wrap(.~HUC_2, ncol=5, labeller = labeller(names=basin.Names))+
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) 
```

###DOC by lake type, size, basin
```{r}
#DOC across lakes with data
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2007 %>%
               filter(LAKE_ORIGIN %in% c("MAN-MADE","NATURAL")),
                                  aes(LON_DD,LAT_DD,
                                   fill=log10(DOC),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Terrain", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#Chlorophyll in lakes of different size classes
#Lakes of different size classes
nla2007 %>%
  select(SIZE_CLASS, DOC, LAKE_ORIGIN) %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(DOC), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  ylab("Log10 DOC")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)+
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) 



#DOC in lakes of different major basins
nla2007 %>%
  select(HUC_2, DOC,LAKE_ORIGIN) %>%
  filter(!HUC_2 %in% c(""))%>%
  mutate(HUC_2=factor(HUC_2,
                      labels=c("Missouri","Ark-Red-White", "Rio Grande",
                 "Upper Colorado","Lower Colorado","Great Basin","Pacific Northwest",
                 "California"))) %>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(DOC), fill=LAKE_ORIGIN))+
  geom_boxplot()+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  ylab("Log10 DOC")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)+
  facet_wrap(.~HUC_2, ncol=5)

```

### TN by lake type, size, basin
```{r}
#TN in lakes of different major basins
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2007 %>%
               filter(LAKE_ORIGIN %in% c("MAN-MADE","NATURAL")),
                                  aes(LON_DD,LAT_DD,
                                   fill=log10(NTL),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Blues", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#TN in lakes of different size classes
#Lakes of different size classes
nla2007 %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(NTL), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  ylab("Log10 TN")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Blues", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=6)+
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) 

#TN in lakes of different major basins
nla2007 %>%
  select(HUC_2, NTL,LAKE_ORIGIN) %>%
  filter(!HUC_2 %in% c(""))%>%
  mutate(HUC_2=factor(HUC_2,
                      labels=c("Missouri","Ark-Red-White", "Rio Grande",
                 "Upper Colorado","Lower Colorado","Great Basin","Pacific Northwest",
                 "California"))) %>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(NTL), fill=LAKE_ORIGIN))+
  geom_boxplot()+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 TN")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Blues", rev=TRUE)+
  facet_wrap(.~HUC_2, ncol=5)
```

###TP by lake type, size, basin
```{r}
#TP in lakes of different major basins
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2007 %>%
               filter(LAKE_ORIGIN %in% c("MAN-MADE","NATURAL")),
                                  aes(LON_DD,LAT_DD,
                                   fill=log10(PTL),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Purple", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#TP in lakes of different size classes
nla2007 %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(PTL), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  ylab("Log10 TP")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Purple", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=6)+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) 

#TP in lakes of different size classes
nla2007 %>%
  ggplot(aes(x=DEPTHMAX, y=log10(PTL), fill=LAKE_ORIGIN))+
  geom_point(shape=21, size=3)+
  ylab("Log10 TP")+
  xlab("Max depth (m)")+
  scale_fill_discrete_sequential(palette = "Purple", rev=TRUE)


#TP in lakes of different major basins
nla2007 %>%
  select(HUC_2, PTL,LAKE_ORIGIN) %>%
  filter(!HUC_2 %in% c(""))%>%
  mutate(HUC_2=factor(HUC_2,
                      labels=c("Missouri","Ark-Red-White", "Rio Grande",
                 "Upper Colorado","Lower Colorado","Great Basin","Pacific Northwest",
                 "California"))) %>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(PTL), fill=LAKE_ORIGIN))+
  geom_boxplot()+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  ylab("Log10 TP")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Purple", rev=TRUE)+
  facet_wrap(.~HUC_2, ncol=6)
```



#Maps - 2012 data
##Basic distribution, size, and lake types
```{r}
#Map 2012 western NLA sites

northamerica<- ne_countries(country = "united states of america", returnclass= "sf")
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-130, -100),
           ylim = c(30, 50)) +
  geom_point(data=nla2012,aes(LON_DD83,LAT_DD83,fill=MAJ_BASIN),
             shape=21, color="black", size=3)+
  scale_fill_discrete_diverging(rev=TRUE,
                                 labels=basinNames)+
  ggtitle("Western US NLA lakes")+
  theme_bw()

# ggplot(data = northamerica) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(
#     xlim = c(-130, -100),
#     ylim = c(30, 50)) +
#   geom_point(data=nla2012,aes(LON_DD83,LAT_DD83,fill=SIZE_CLASS),
#              shape=21, color="black")+
#   scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)


#How many of these lakes are reservoirs?  
nla2012 %>%
  select(MAJ_BASIN,LAKE_ORIGIN) %>%
  filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL"))%>%
  group_by(MAJ_BASIN,LAKE_ORIGIN)%>%
  mutate(num_lakes=n()) %>%
  distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=num_lakes, fill=MAJ_BASIN))+
  geom_bar(stat="identity",  color="black")+
  ylab("Number of lakes")+
  xlab("Lake type")+
  facet_wrap(.~MAJ_BASIN, scales="free_y")+
  scale_fill_discrete_diverging(rev=TRUE,
                                labels=basinNames)+
  ggtitle("Lake origin by major basin")

#Lakes of different size classes
nla2012 %>%
  select(SIZE_CLASS) %>%
  filter(!SIZE_CLASS=="")%>%
  # pivot_longer(-1) %>%
  group_by(SIZE_CLASS) %>%
  mutate(num_lakes=n()) %>%
  distinct() %>% #remove duplicates
  ggplot(aes(x=SIZE_CLASS, y=num_lakes, fill=SIZE_CLASS))+
  geom_bar(stat="identity",  color="black")+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Number of lakes")+
  xlab("Size class")+
  # geom_text(aes(label=load_calc_possible_n), vjust=-1) +
  theme(legend.position="none")+
  ggtitle("Number of lakes by size class")


#How many fewer sites do we have if we omit the lakes without DOC, TN, TP data?
nla2012_bare <- nla2012 %>%
  drop_na(CHLX_RESULT)
```

##Chemistry
###Chlorophyll by lake type, size, basin
```{r}
#Chlorophyll across lakes with data
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2012_bare %>%
               filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL")),
                                  aes(LON_DD83,LAT_DD83,
                                   fill=log10(CHLX_RESULT),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Greens", rev=TRUE)+
  scale_shape_manual(values=c(21,23))

#Chlorophyll in lakes of different size classes
#Lakes of different size classes
nla2012_bare %>%
  select(SIZE_CLASS, CHLX_RESULT, LAKE_ORIGIN) %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(CHLX_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 chlorophyll a")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Greens", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)



#Chlorophyll in lakes of different major basins
nla2012_bare %>%
  select(MAJ_BASIN, CHLX_RESULT,LAKE_ORIGIN) %>%
  filter(!MAJ_BASIN %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(CHLX_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 chlorophyll a")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Terrain2", rev=TRUE)+
  facet_wrap(.~MAJ_BASIN, ncol=5)
```

###DOC by lake type, size, basin
```{r}
#DOC across lakes with data
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2012_bare %>%
               filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL")),
                                  aes(LON_DD83,LAT_DD83,
                                   fill=log10(DOC_RESULT),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Terrain", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#Chlorophyll in lakes of different size classes
#Lakes of different size classes
nla2012_bare %>%
  select(SIZE_CLASS, DOC_RESULT, LAKE_ORIGIN) %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(DOC_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 DOC")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)



#DOC in lakes of different major basins
nla2012_bare %>%
  select(MAJ_BASIN, DOC_RESULT,LAKE_ORIGIN) %>%
  filter(!MAJ_BASIN %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(DOC_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 DOC")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Terrain", rev=TRUE)+
  facet_wrap(.~MAJ_BASIN, ncol=5)

```

### TN by lake type, size, basin
```{r}
#TN in lakes of different major basins
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2012_bare %>%
               filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL")),
                                  aes(LON_DD83,LAT_DD83,
                                   fill=log10(NTL_RESULT),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Blues", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#TN in lakes of different size classes
#Lakes of different size classes
nla2012_bare %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(NTL_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 TN")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Blues", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)

#TN in lakes of different major basins
nla2012_bare %>%
  select(MAJ_BASIN, NTL_RESULT,LAKE_ORIGIN) %>%
  filter(!MAJ_BASIN %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(NTL_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
    stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 TN")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Blues", rev=TRUE)+
  facet_wrap(.~MAJ_BASIN, ncol=5)
```

###TP by lake type, size, basin
```{r}
#TP in lakes of different major basins
ggplot(data = northamerica) +
  geom_sf() +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(
    xlim = c(-125, -100),
    ylim = c(30, 50)) +
  geom_point(data=nla2012_bare %>%
               filter(LAKE_ORIGIN %in% c("MAN_MADE","NATURAL")),
                                  aes(LON_DD83,LAT_DD83,
                                   fill=log10(PTL_RESULT),
                                   shape=LAKE_ORIGIN),
             color="black", size=3, alpha=0.9)+
  scale_fill_continuous_sequential(palette = "Purple", rev=TRUE)+
  scale_shape_manual(values=c(21,23))


#TP in lakes of different size classes
nla2012_bare %>%
  filter(!SIZE_CLASS %in% c(""))%>%
  mutate(LAKE_ORIGIN=factor(LAKE_ORIGIN,
                            labels=c("Res.","Natural"))) %>%
  ggplot(aes(x=LAKE_ORIGIN, y=log10(PTL_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
      stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  ylab("Log10 TP")+
  xlab("Lake type")+
  scale_fill_discrete_sequential(palette = "Purple", rev=TRUE)+
  facet_wrap(.~SIZE_CLASS, ncol=5)+
  theme_bw()+
  theme(legend.position="none")

#TP in lakes of different major basins
nla2012_bare %>%
  select(MAJ_BASIN, PTL_RESULT,LAKE_ORIGIN) %>%
  filter(!MAJ_BASIN %in% c(""))%>%
  # pivot_longer(-1) %>%
  # distinct() %>% #remove duplicates
  ggplot(aes(x=LAKE_ORIGIN, y=log10(PTL_RESULT), fill=LAKE_ORIGIN))+
  geom_boxplot()+
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  # scale_fill_manual(values=c("#d62828","#003049"))+
  ylab("Log10 TP")+
  xlab("Major basins")+
  scale_fill_discrete_sequential(palette = "Purple", rev=TRUE)+
  facet_wrap(.~MAJ_BASIN, ncol=5)
```

#Combine 2007 + 2012 datsets
```{r}
glimpse(nla2007)
glimpse(nla2012)

#RENAME some columns for easier merging
nla2012_combine<-nla2012 %>%
  rename(NTL=NTL_RESULT,
         PTL=PTL_RESULT,
         TURB=TURB_RESULT,
         DOC=DOC_RESULT,
         NITRATE_N=NITRATE_N_RESULT,
         LAT_DD=LAT_DD83,
         LON_DD=LON_DD83,
         NO3_N=NITRATE_N_RESULT,
         SITEID_12=SITE_ID) %>%
  mutate(DATE_COL=mdy(DATE_COL),
         YEAR=year(DATE_COL)) 

nla2007_combine<-nla2007 %>%
  rename(HUC2=HUC_2,
         HUC8=HUC_8,
         SIZE_CLASS_07=SIZE_CLASS,
         NO3N=NO3N_PPM, 
         SITEID_07=SITE_ID) %>%
  mutate(LAKE_ORIGIN=case_when(
  LAKE_ORIGIN == "MAN-MADE" ~ "MAN_MADE",
  LAKE_ORIGIN == "NATURAL" ~ "NATURAL",
  TRUE ~ "other")) %>%
  select(-LAKEAREA) %>%
  mutate(DATE_COL=mdy(DATE_COL),
         YEAR=year(DATE_COL)) %>%
    select(-PTL_2, -NTL_2, -DOC_2,
         -TURB_2, -CHLA_2)

glimpse(nla2007_combine)
glimpse(nla2012_combine)

# nla_all <- left_join(nla2007_combine, nla2012_combine, by=c("SITE_ID"="SITEID_07"))
nla2007_combine_smol <- nla2007_combine %>%
  select(SITEID_07, DATE_COL)
nla2012_combine_smol <- nla2012_combine %>%
  select(SITEID_07, SITEID_12, DATE_COL) %>%
  drop_na(DATE_COL)

#How many lakes were sampled in 2007 AND 2012? 
left_join(nla2012_combine_smol,nla2007_combine_smol, by="SITEID_07") %>%
  drop_na() %>%
  n_distinct()

test_join<-left_join(nla2012_combine,nla2007_combine, by=c("SITEID_07"))
```